---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(survival)
library(ggfortify)
library(ggthemes)
library(RColorBrewer)
```

```{r}
path <- "Data/veteran.csv"
df <- read_csv(path)
df
```

```{r}
df %>% map(., ~ typeof(.)) %>% cbind(type = .)
df %>% mosaic::inspect()
```

```{r}
km <- survfit(Surv(time, status)~1., data = df)
km_table <- ggplot2::fortify(km)
km_table %>% tbl_df
```
#### Confirm how to calc Kaplan Meier survival curve and standard error
```{r}
km_table %>% mutate(r = (n.risk-n.event)/n.risk) %>% 
  mutate(my_surv = cumprod(r)) %>% 
  mutate(temp = n.event/((n.risk)*(n.risk-n.event))) %>% 
  mutate(my_se = sqrt(cumsum(temp))) %>% select(-temp) %>% tbl_df

```

#### draw survival curv
```{r}
p1 <- km_table %>% ggplot(aes(x = time, y = surv))+ geom_line() + 
  geom_point() + 
  geom_ribbon(aes(ymin=lower, ymax = upper), alpha = 0.3)+
  theme_bw(base_size = 16, base_family = "Palatino") +
  ylab("survival") + xlab("time (/months)") 

nodes <- 8
time_nodes_ <- min(km_table$time) + (max(km_table$time) - min(km_table$time))/nodes * 1:nodes
time_nodes  <- km_table$time[sapply(time_nodes_, function(x) which.min(abs(km_table$time - x)))]
p2 <- km_table %>% filter(time %in% time_nodes) %>% ggplot(aes(x = time))+
  geom_text(aes(label = n.risk, y= 1)) + 
  theme_bw(base_family = "Palatino", base_size = 16) + 
  theme(axis.text.y= element_blank())

egg::ggarrange(p1, p2, nrow =2, heights = c(0.8, 0.2))
```

```{r}
km_fit_trt <- survfit(Surv(time, status)~trt, data = df)
km_fit_trt %>% fortify -> trt_table
p <- km_fit_trt %>% autoplot
p$theme <- theme_bw(base_family = "Times", base_size = 16)
p
```

```{r}
age_cuts <- df$age %>% quantile(prob = c(0.333, 0.667))
brks <- c(0, age_cuts, max(df$age))
df$age_c <- df$age %>% cut(breaks = brks, labels = str_c("[", brks[1:length(brks)-1], "_", lead(brks)[1:length(brks)-1], ")")) 

km_fit_age <- survfit(Surv(time, status)~age_c, data =df)
km_age_table <- fortify(km_fit_age)
p_age <- km_fit_age %>% autoplot
p_age$theme <- theme_bw(base_family = "Times", base_size = 16) + 
  theme(legend.position = c(0.85, 0.85), 
        legend.justification = c(0.85, 0.85),
        axis.title.x = element_blank())

p_age_text <- km_age_table %>% group_by(strata) %>% 
  filter(time %in% time[sapply(time_nodes_, function(x)which.min(abs(time - x) ))]) %>% 
  ungroup() %>% arrange(strata, time) %>% 
  ggplot(aes(x = time, y = strata))+ 
  geom_text(aes(label = n.risk, color = strata)) +
  theme_bw(base_size = 16, base_family = "Times")+
  theme(legend.position = "none")

egg::ggarrange(p_age, p_age_text, heights = c(0.75, 0.25))
```
#### Cox ph
```{r}
cox <- coxph(Surv(time, status) ~ trt + celltype + age + karno + diagtime + prior, data = df)
cox
```
### Porportional hazard
#### Complementary log-log plot
$$
 \log(-\log{S(t)}) = \log(-\log{S_{0}(t)}) + (\beta_{0} + \beta_{1}x_{1}+\beta_{2}x_{2}+\cdots+\beta_{n}x_{n})\\
 (-\log{S(t)=H(t)})
$$

### Log-log
$\log(y)$ vs $\log\{-\log(S(t))\}$ 

```{r}
km_age_table %>% mutate(ll = log(-log(surv))) %>% 
  mutate(logt = log(time)) %>% 
  ggplot(aes(x = logt, y = ll)) + 
  geom_point(aes(color = strata, fill = strata)) +
  theme_linedraw(base_size = 16, base_family = "Times")

```

#### For 2 group case
```{r}
age_med <- df$age %>% quantile %>% `[`(3)
df$age_c2 <- df$age %>% cut(breaks = c(0, age_med, max(df$age)), labels = c("younger", "elder"), include.lowest = T, right = F)
km_age_table2g <- survfit(Surv(time, status)~age_c2, data = df) %>% fortify
km_age_table2g%>% mutate(ll = log(-log(surv))) %>% 
  mutate(logt = log(time)) %>% 
  ggplot(aes(x = logt, y = ll)) + 
  geom_point(aes(color = strata, fill = strata)) +
  theme_linedraw(base_size = 16, base_family = "Times")
```

#### Log rank 
rh = 0 => log-rank, rh =1 => Gehan-Wilcoxon
```{r}
survdiff(Surv(time, status)~age_c2, data = df)
survdiff(Surv(time, status)~karno, data = df)
survdiff(Surv(time, status)~trt, data = df)
```

#### Schoenfeld residuals
Cox zph caclulates Schoenfeld residual.
[Vignette:timedep](https://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf)
Using this residuals we can test hazard ratio under null hypothesis "the coefficient satisfies the proportional hazard assumption"
```{r}
ggplot(data = df, aes(x = time, y = karno)) + geom_line()
zp <- cox.zph(cox, transform= function(time) log(time +20))
zp
```

```{r}
zp %>% plot
```

So, for trt, celltypelarge, celltypesmallcell, celltypesquamous and diagtime, we cannot strongly insist on non-proportionality.
But other features, in other words, age, karno and prior, non-proportionality may be behind the scene.
+ Multiple hazard functions of which dependencies on time differ each other.
+ Time-dependent coefficients
+ Log-linearity is wrong

How to deal with this problem,
+ Use different base line hazards for different subgroupes
+ Assume time-dependent coefficient
+ Add non-linear effect to linear term in cox model
```{r}
vet2 <- survSplit(Surv(time, status) ~ ., data= veteran, cut=c(90, 180), episode= "tgroup", id="id")
vet2 %>% tbl_df
```

```{r}
vfit2 <- coxph(Surv(tstart, time, status) ~ trt + prior + karno:strata(tgroup), data=vet2)
vfit2
```

```{r}
cox.zph(vfit2)
```